---
title: "Tx_analysis"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = FALSE} 
#Load libraries
library(ggplot2)
library(purrr)
library(tidyverse)
library(beeswarm)
library(readr)
library(tidyr)
library(Hmisc)
library(plyr)
library(RColorBrewer)
library(reshape2)
library(dplyr)
library(stringr)
library(reshape2)
library(ggthemes)
library(pander)
library(ggpubr)
```

## Make plotting themes
```{r}
man_fill <- scale_fill_manual(values = c("N2" = "#FCE85C", 
                                         "CB428" = "#44A49C",
                                         "RNAi_ama1" = "#CB297B",
                                         "RNAi_dpy27" = "#57A56F", 
                                         "RNAi_sdc2" = "#FF9300", 
                                         "SEA-12" = "#EE220C", 
                                         "NA" = "#A9A9A9",
                                         "RNAi_rluc" = "#A9A9A9" ,
                                         "RNAi_rlac" = "#A9A9A9",                                
                                         "RNAi" = "#A9A9A9"  ))
  
  
man_color <-     scale_color_manual(values = c("N2" = "#FCE85C", 
                                         "CB428" = "#44A49C",
                                         "RNAi_ama1" = "#CB297B",
                                         "RNAi_dpy27" = "#57A56F", 
                                         "RNAi_sdc2" = "#FF9300", 
                                         "SEA-12" = "#EE220C", 
                                         "NA" = "#A9A9A9",
                                         "RNAi_rluc" = "#A9A9A9" ,
                                         "RNAi_rlac" = "#A9A9A9",                                
                                         "RNAi" = "#A9A9A9"))
```

#Functions
```{r}

n_fun <- function(x){
  return(data.frame(y = 0.95*30,
                    label = length(x)))
}
```

#Import the data
###-Read the CSV file
```{r, include=FALSE}
smFISH <- read_csv("/Users/lbreima/Documents/PhD_local/smFISH_analysis/191213_tx_counts_from_exon/master_with_more_columns_threshold-3.csv")
```

###-Check that the column types are right
```{r, echo = FALSE}
spec(smFISH)
```

###-Report problems
```{r, echo = FALSE}
problems(smFISH)
```

###-Make a new colum for the strain name and plot unique strain names
```{r, echo = FALSE}
new_smfish <- smFISH %>%
  mutate(strain = gsub("(.+)\\_.+","\\1",`filename`))

unique(new_smfish$strain, incomparables = FALSE)
```

###-Remove the # from the smFISH counts (because I couldnt plot with them?!?)
```{r, echo = FALSE}
colnames(new_smfish)[colnames(new_smfish)=="#c0_smfish"] <- "c0_smfish"
colnames(new_smfish)[colnames(new_smfish)=="#c1_smfish"] <-  "c1_smfish"
colnames(new_smfish)[colnames(new_smfish)=="#c2_smfish"] <- "c2_smfish"
colnames(new_smfish)[colnames(new_smfish)=="#c0_smfish_adj"] <- "c0_smfish_adj"
colnames(new_smfish)[colnames(new_smfish)=="#c1_smfish_adj"] <- "c1_smfish_adj"
colnames(new_smfish)[colnames(new_smfish)=="#c2_smfish_adj"] <- "c2_smfish_adj"
colnames(new_smfish)[colnames(new_smfish)=="#nuclei"] <- "manual_nuclei"
colnames(new_smfish)[colnames(new_smfish)=="#nucs_predicted"] <- "nucs_predicted"
colnames(new_smfish)[colnames(new_smfish)=="X.c0_smfish"] <- "c0_smfish"
colnames(new_smfish)[colnames(new_smfish)=="X.c1_smfish"] <-  "c1_smfish"
colnames(new_smfish)[colnames(new_smfish)=="X.c2_smfish"] <-"c2_smfish"
colnames(new_smfish)[colnames(new_smfish)=="X.nuclei"] <- "manual_nuclei"
colnames(new_smfish)[colnames(new_smfish)=="X.nucs_predicted"] <-  "nucs_predicted"
colnames(new_smfish)[colnames(new_smfish)==" X.channels"] <- "channels"
colnames(new_smfish)[colnames(new_smfish)=="X.c0_smfish_adj"] <- "c0_smfish_adj"
colnames(new_smfish)[colnames(new_smfish)=="X.c1_smfish_adj"] <- "c1_smfish_adj"
colnames(new_smfish)[colnames(new_smfish)=="X.c2_smfish_adj"] <- "c2_smfish_adj"
```

###-Make a new colum with the date of the experiment
```{r, echo = FALSE}
colnames(new_smfish)[colnames(new_smfish)=="original.filename"] <- "original_filename"

new_new_smfish <- new_smfish %>%
     mutate(exp.date = gsub("(.?)\\_.+","\\1",original_filename))
```

# Change some emrbyo stages by hand that got wrong label
```{r}
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] == "RNAi_dpy27_1155_cropped_5104.tif"] <- 5
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] == "N2_293_cropped_728.tif"] <- 1
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] == "N2_362_cropped_1251.tif"] <- 1
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="SEA-12_324_cropped_166.tif"] <- 1
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="RNAi_dpy27_3741_cropped_0.tif"] <- 2
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="RNAi_dpy27_3743_cropped_1.tif"] <- 0
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="N2_3684_cropped_0.tif"] <- 3
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="N2_3690_cropped_0.tif"] <- 2
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="N2_3709_cropped_0.tif"] <- 2
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="N2_3596_cropped_0.tif"] <- 1
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="N2_3601_cropped_0.tif"] <- 1
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="N2_3609_cropped_0.tif"] <- 1
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="N2_3668_cropped_0.tif"] <- 1
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="N2_3684_cropped_1.tif"] <- 1
new_new_smfish[,'predicted_bin'][new_new_smfish[,'cropped_image_file'] =="N2_3708_cropped_0.tif"] <- 1

```

# Filter broken embryos
```{r}

new_new_smfish<-new_new_smfish[!(new_new_smfish$cropped_image_file=="RNAi_dpy27_1761_cropped_6375.tif" 
                      | new_new_smfish$cropped_image_file=="RNAi_dpy27_1167_cropped_5125.tif"
                      | new_new_smfish$cropped_image_file=="RNAi_dpy27_398_cropped_2898.tif"
                      | new_new_smfish$cropped_image_file=="RNAi_dpy27_485_cropped_3037.tif"
                      | new_new_smfish$cropped_image_file=="RNAi_dpy27_1790_cropped_6660.tif"
                      | new_new_smfish$cropped_image_file=="RNAi_sdc2_1302_cropped_5665.tif"
                       | new_new_smfish$cropped_image_file=="N2_1541_cropped_2243.tif"
                       | new_new_smfish$cropped_image_file=="N2_1266_cropped_2385.tif"
                      | new_new_smfish$cropped_image_file== "N2_1025_cropped_3847.tif"
                     | new_new_smfish$cropped_image_file==  "N2_1025_cropped_3846.tif"
                     | new_new_smfish$cropped_image_file==  "RNAi_sdc2_1311_cropped_5675.tif"
                      | new_new_smfish$cropped_image_file==  "RNAi_dpy27_405_cropped_2905.tif" 
                      ),]
```

### Combine stages from predicted_bin and stage_bin
```{r}
new_new_smfish$stage_bin <- replace(new_new_smfish$stage_bin, which(new_new_smfish$stage_bin < 0), NA)
new_new_smfish$predicted_bin <- replace(new_new_smfish$predicted_bin, which(new_new_smfish$predicted_bin < 0), NA)

new_new_smfish <- new_new_smfish %>% mutate(combined_stage = coalesce(stage_bin,predicted_bin))
```

###-Make a new colum with stage classification
```{r, echo = FALSE}
new_new_smfish <- new_new_smfish %>%
     mutate(stage.class = ifelse( combined_stage==0, "1-4", 
                            ifelse ( combined_stage==1, "5-49", 
                              ifelse ( combined_stage==2, "50-99",
                                ifelse ( combined_stage==3, "100-149",  
                                  ifelse ( combined_stage==4, "150-534",        
                                    ifelse ( combined_stage==5, "535-558",NA )))))))
                                      

new_new_smfish <- new_new_smfish %>%
     mutate(prediction = ifelse( combined_stage==0, 2, 
                          ifelse ( combined_stage==1, 27, 
                            ifelse ( combined_stage==2, 70,
                              ifelse ( combined_stage==3, 125,  
                                ifelse ( combined_stage==4, 246,        
                                  ifelse ( combined_stage==5, 547, NA )))))))
```

## Make a table were all smFISH counts are in one column
```{r}
# tx_count_adj is the number of transcripts at the transcription site
# tx_count is the number of transcription sites per embryo

c0 <-select(new_new_smfish, c0_smfish,c0_smfish_adj, c0_type, channel_1_counts, sum_of_intensity_1, filename:status, cropped_image_file, is_dapi_stack:is_valid_final,is_male_batch:num_z_planes, stage_bin, predicted_bin,stage.class ,prediction, strain )

c1 <-select(new_new_smfish, c1_smfish,c1_smfish_adj, c1_type, channel_2_counts, sum_of_intensity_2, filename:status, cropped_image_file, is_dapi_stack:is_valid_final,is_male_batch:num_z_planes, stage_bin,predicted_bin, stage.class, prediction ,strain )

c2 <-select(new_new_smfish, c2_smfish,c2_smfish_adj, c2_type, channel_3_counts, sum_of_intensity_3, filename:status, cropped_image_file, is_dapi_stack:is_valid_final,is_male_batch:num_z_planes, stage_bin,predicted_bin, stage.class, prediction ,strain )

colnames(c0)[colnames(c0)=="c0_smfish"] <- "smfish_count"
colnames(c0)[colnames(c0)=="c0_smfish_adj"] <- "smfish_count_adj"
colnames(c0)[colnames(c0)=="channel_1_counts"] <- "tx_count"
colnames(c0)[colnames(c0)=="sum_of_intensity_1"] <- "tx_count_adj"

colnames(c1)[colnames(c1)=="c1_smfish"] <- "smfish_count"
colnames(c1)[colnames(c1)=="c1_smfish_adj"] <- "smfish_count_adj"
colnames(c1)[colnames(c1)=="channel_2_counts"] <- "tx_count"
colnames(c1)[colnames(c1)=="sum_of_intensity_2"] <- "tx_count_adj"

colnames(c2)[colnames(c2)=="c2_smfish"] <- "smfish_count"
colnames(c2)[colnames(c2)=="c2_smfish_adj"] <- "smfish_count_adj"
colnames(c2)[colnames(c2)=="channel_3_counts"] <- "tx_count"
colnames(c2)[colnames(c2)=="sum_of_intensity_3"] <- "tx_count_adj"

colnames(c0)[colnames(c0)=="c0_type"] <- "gene"
colnames(c1)[colnames(c1)=="c1_type"] <- "gene"
colnames(c2)[colnames(c2)=="c2_type"] <- "gene"

allgenes <- bind_rows(c0, c1)
allgenes <- bind_rows(allgenes, c2)
```

Make collumn for transcripts per TSS
```{r}
allgenes <- allgenes %>%
     mutate(ave_tx_int =  tx_count_adj/tx_count)
```

# Filters
```{r}
allgenes_true_counts <- filter(allgenes,  status ==1, prediction != "NA", is_valid_final !=0, is_male_batch !=1, is_z_cropped != 1, num_z_planes >60 ,  tx_count <1116 , tx_count >0 )


strain.names <- unique(allgenes_true_counts$strain)
gene.names <- unique(allgenes_true_counts$gene)

allgenes_length <- ddply(.data=allgenes_true_counts, 
                  .(strain, gene),
                 summarise, 
                 n=paste(length(tx_count)))


allgenes_true_counts$strain <- factor(allgenes_true_counts$strain, levels =c("N2", "RNAi_dpy27", "MK4", "CB428", "RNAi_set1", "RNAi_set4" , "RNAi_sdc2", "RNAi_ama1", "SEA-12", "RNAi_rluc", "RNAi_rlac", "RNAi", "RNAi-pop1"))

allgenes_true_counts$stage.class <- factor(allgenes_true_counts$stage.class, levels =c("1-4", "5-49", "50-99", "100-149", "150-534", "535-558"))

```

# Check some specfific data points
```{r}
young_embryos_too_high <- filter(allgenes_true_counts, stage.class == "1-4", tx_count > 8)

allgenes_true_counts$multi_tx <- allgenes_true_counts$tx_count_adj / allgenes_true_counts$prediction
```

# Dotplot for DPY-27
```{r}
func_dot = function(k,n) {
  allgenes_filtered <- s %>%
    filter(gene == k) %>%
    ggdotplot(x= "stage.class", 
              y = "tx_count",
              add = c("boxplot","jitter"),
              add.params = list(width = 0.5),
              color = "strain",
              fill ="strain",
              error.plot = "crossbar",
              stackdir = "center",
              stackratio = 0.2,
              size = 0.5)+
          man_fill +
          man_color +
          labs(x = "binned nuclei") +
    stat_summary(fun.data = n_fun, geom = "text",
               aes(group = strain),
               hjust = 0.5 ,
               color = "black",
             vjust = 3,
               position = position_dodge(0.7)) +
    
  scale_y_continuous(sec.axis = dup_axis(label = NULL, 
                                       name = NULL),
                    expand = expand_scale(mult = c(0.1, 0)),
             breaks = pretty(c(0,100), n = 5), 
                      limits = c(0,40))+
          ggtitle('Number of transcription sites per embryo, gene:',paste(k))
               }

s <- allgenes_true_counts %>%
  filter(strain== "N2" | strain== "RNAi_dpy27" )
 
strain.names <- unique(s$strain)
gene.names <- unique(s$gene)
 
 
 
  for (k in gene.names){
                        n <- allgenes_length %>%
                          filter(gene == k) %>%
                          pull(n)
    hist2 <- func_dot(k,n)
    ggsave(hist2, file = paste0(k, "_tx_sites_dpy27.pdf"), scale = 1)
    print(hist2)
  }
```


# Dotplot for DPY-27 transcripts
```{r}
func_dot = function(k,n) {
  allgenes_filtered <- s %>%
    filter(gene == k) %>%
    ggdotplot(x= "stage.class", 
              y = "tx_count_adj",
              add = c("boxplot","jitter"),
              add.params = list(width = 0.5),
              color = "strain",
              fill ="strain",
              error.plot = "crossbar",
              stackdir = "center",
              stackratio = 0.2,
              size = 0.5)+
          man_fill +
          man_color +
          labs(x = "binned nuclei") +
    stat_summary(fun.data = n_fun, geom = "text",
               aes(group = strain),
               hjust = 0.5 ,
               color = "black",
             vjust = 3,
               position = position_dodge(0.7)) +
  #expand_limits(y = 0) +
  scale_y_continuous(sec.axis = dup_axis(label = NULL, 
                                       name = NULL),
                    expand = expand_scale(mult = c(0.1, 0)),
             breaks = pretty(c(0,150), n = 5), 
                      limits = c(0,150))+
          ggtitle('Number of nascent transcripts per emrbyo, gene:',paste(k))
               }

s <- allgenes_true_counts %>%
  filter(strain== "N2" | strain== "RNAi_dpy27" )

 
 
strain.names <- unique(s$strain)
gene.names <- unique(s$gene)
 
 
 
  for (k in gene.names){
                        n <- allgenes_length %>%
                          filter(gene == k) %>%
                          pull(n)
    hist2 <- func_dot(k,n)
    ggsave(hist2, file = paste0(k, "_txs_dpy27.pdf"), scale = 1)
    print(hist2)
  }
```

##Dotplot for DPY-27 transcripts per TSS
```{r}

func_dot = function(k,n) {
  allgenes_filtered <- s %>%
    filter(gene == k) %>%
    ggdotplot(x= "stage.class", 
              y = "ave_tx_int",
              add = c("boxplot","jitter"),
              add.params = list(width = 0.5),
              color = "strain",
              fill ="strain",
              error.plot = "crossbar",
              stackdir = "center",
              stackratio = 0.2,
              binwidth = 0.5, 
              size = 0.5)+
          man_fill +
          man_color +
          labs(x = "binned nuclei") +
    stat_summary(fun.data = n_fun, geom = "text",
               aes(group = strain),
               hjust = 0.5 ,
               color = "black",
             vjust = 3,
               position = position_dodge(0.7)) +
  #expand_limits(y = 0) +
  scale_y_continuous(sec.axis = dup_axis(label = NULL, 
                                       name = NULL),
                    expand = expand_scale(mult = c(0.1, 0)),
             breaks = pretty(c(0,50), n = 5), 
                      limits = c(0,50))+
          ggtitle('Number of nascent transcripts per TSS, gene:',paste(k))
               }

s <- allgenes_true_counts %>%
  filter(strain== "N2" | strain== "RNAi_dpy27" )

 
 
strain.names <- unique(s$strain)
gene.names <- unique(s$gene)
 
 
 
  for (k in gene.names){
                        n <- allgenes_length %>%
                          filter(gene == k) %>%
                          pull(n)
    hist2 <- func_dot(k,n)
    ggsave(hist2, file = paste0(k, "_txs_2_dpy27.pdf"), scale = 1)
    print(hist2)
  }

```



